Response to decision letter

### 2023-11-11 submission/review2.Rmd
Edits from second review.

---
title: "Response to decision letter"
output: 
  pdf_document:
    keep_md: yes
date: "2023-07-20"
---

@@

**Thank you for your decision! We modified Figures 1 and 4.**

Figure 5 is too large and detailed: it could never be published as an A4. 
I also wonder if all figures are necessary: for example, Figures 5 and 6 could be moved to the Supplementary; there is no limit in the number of figures in Software papers but we typically publish 4 figures per paper. Please consider these recommendations as you revise the manuscript.

**We moved figures 5 and 6 to supplementary materials.**

@@

**Noted. We highlighted changes using latex and submitted the PDF file.**

@@

**We added the DOI of our Dryad package containing all Supplementary Material on L690 (manuscript-differences.pdf).**

@@

**Done on L700 (manuscript-differences.pdf).**

@@

**Checked.**

@@

Thanks you for your decision! We addressed all the following comments.

@@

Rephrased to "a phylogenetic software that provides services"

@@

By default it uses OpenTree's tree topology, ig the user does not provide a topology that they want. We specified this as "to corresponding nodes on a tree topology, either extracted from Open Tree of Life's synthetic phylogeny or one provided by the user"

@@

Thanks for pointing this out, we rephrased to  "Because the results of any […], we highlight that […]"

@@

Yes, we changed to "DateLife can help to increase [...]"

@@

@@

@@

Thanks for pointing it out! We changed it to:

"BLADJ is an agnostic way to assign ages to nodes with no available genetic or branch length data, as it does not require initial diversification rates to draw from a branch length distribution"

An alternative to BLADJ to date tree topologies in the absence of initial branch lengths that is common practice in the literature is to use a birth-death model to draw branch lengths [@Jetz2012; @rabosky2018inverse; @smith2018constructing]. In addition to the initial tree topology and nodes with age data, these methods require initial values of speciation and extinction rate parameters provided by the user. DateLife implements this approach with MrBayes [@Huelsenbeck2001; @Ronquist2003], using nodes with published age data as calibration priors on nodes of a tree topology with no branch lengths, a simple birth-death model with speciation and extinction rate parameters that are provided by the user, and no genetic data. However, BLADJ is the default option in DateLife, as it does not require any information on diversification rates for the phylogenetic sample to draw from a branch length distribution.

@@

@@

@@

@@

@@

We streamlined the figure by removing text in the figure and adding it to the figure caption. We considered removing the tables, but we decided to keep them as they are important for the explanation of the workflow. We hope this new version of the figure is considerably clearer now.

@@

@@

@@

@@

>>>>

---
title: "Response to decision letter"
output: 
  pdf_document:
    keep_md: yes
date: "2023-11-07"
---

@@

**Thank you for your decision! We modified Figures 1 and 4, and consolidated into a single one. They are more streamlined now and usage of color was reduced; we followed all suggestions from AE.**

Figure 5 is too large and detailed: it could never be published as an A4.

**We reduced the number of tips shown in Fig 5 and moved the tree with all tips to supplementary data. Fig 5 is now Fig 3**

I also wonder if all figures are necessary: for example, Figures 5 and 6 could be moved to the Supplementary; there is no limit in the number of figures in Software papers but we typically publish 4 figures per paper. Please consider these recommendations as you revise the manuscript.

**We have now 4 figures in our paper. We achieved this by consolidating fig 1 and 3 into one single figure (now fig 2) and we moved figures 4 and 5 to supplementary material. We would like to request keeping Figure 6, which is now Figure 4 in the new submission, as we believe it adds good visual information for the cross validation analysis.**

@@

**Noted. We highlighted changes using latex and submitted the PDF file with track changes names manuscript-differences.pdf.**

@@

**We added the DOI of our Dryad package containing all Supplementary Material at the en dof the manuscript (https://datadryad.org/stash/dataset/doi:10.5061/dryad.cnp5hqc6w).**

@@

**Added in "Funding" section, after "Supplementary Material".**

@@

**Checked and implemented.**

@@

**Thanks you for your decision! We addressed all your comments as follows.**

@@

**Rephrased to "a phylogenetic software that provides services"**

@@

**By default it uses OpenTree's tree topology, ig the user does not provide a topology that they want. We specified this as "to corresponding nodes on a tree topology, either extracted from Open Tree of Life's synthetic phylogeny or one provided by the user"**

@@

**Thanks for pointing this out, we rephrased to  "Because the results of any […], we highlight that […]"**

@@

**Yes, we changed to "DateLife can help to increase [...]"**

@@

**We address this on the manuscript, section "Searching a chronogram database", second paragraph (L 126-132). Briefly, it depends on the taxonomic scope that the researcher is looking for. The chronogram database encompasses close to 100 k (99 474) different species. The chronograms in datelife’s database encompass almost the same amount of tips with phylogenetic information in OpenTree’s synthetic tree (112 669), in which Chordata and Embryophyta are nearly fully sampled. In contrast, Bacteria, Fungi, Nematoda, and Insecta, present a large gap between the estimated number of species and what has phylogenetic information in OpenTree’s synthetic tree. We would expect that users working with the former groups will always get results from datelife. If none of the user’s species are found, datelife will return a message saying that there are no chronograms containing any of those species in the database. It will also encourage users to submit chronograms to a public database, so that their taxon of interest can be included in future searches.**

@@

**Yes! They can add any chronograms they want to the database, or by providing an initial chronogram, node ages and tip names will be used as input for the analysis.**

@@

**Thanks for pointing it out! We changed it to:**

**"BLADJ is an agnostic way to assign ages to nodes with no available genetic or branch length data, as it does not require initial diversification rates to draw from a branch length distribution"**

**An alternative to BLADJ to date tree topologies in the absence of initial branch lengths that is common practice in the literature is to use a birth-death model to draw branch lengths [@Jetz2012; @rabosky2018inverse; @smith2018constructing]. In addition to the initial tree topology and nodes with age data, these methods require initial values of speciation and extinction rate parameters provided by the user. DateLife implements this approach with MrBayes [@Huelsenbeck2001; @Ronquist2003], using nodes with published age data as calibration priors on nodes of a tree topology with no branch lengths, a simple birth-death model with speciation and extinction rate parameters that are provided by the user, and no genetic data. However, BLADJ is the default option in DateLife, as it does not require any information on diversification rates for the phylogenetic sample to draw from a branch length distribution.**

@@

**In this instance we are referring to branch lengths from phylogenetic trees that have not been dated yet. We modified the phrasing to make it more clear.**

@@

**Yes! we clarified that, too.**

@@

**We removed it.**

@@

**We simplified all figures and reduced the number of final figures in the main manuscript.**

@@

@@

**We consolidated Fig 1 and 3 and streamlined the figure by removing text in the figure and adding it to the figure caption. We removed the tables ad only reference them. We hope this new version of the figure is considerably clearer now.**

@@

**We reduced the number of tips shown in Figure 5 and sent the full figure (as well as Figure 4) to supplementary material. Figure 5 is now Figure 3.**

@@

**Thanks so much!**

@@

**We greatly appreciate your consideration of our resubmission.**

**Luna Sanchez-Reyes on behalf of all coauthors.**
